Convergent Evolution During Local Adaptation to Patchy Landscapes

### 2014-11-27 patchy-selection-paper.tex
hwew

Taking these approximations at face value, 
and assuming both rates are small, 
the time until the first of the two patches is colonized by the $B$
allele will be approximately exponentially distributed with mean $1/(2
\mutrate)$.
Following this, the time until the second patch is subsequently colonized 
(via either migration or new mutation) 
will be exponentially distributed with mean $1/(\mutrate+\migrate)$.
Both these rates scale linearly with population density ($\rho$) 
and the probability of establishment of a rare allele ($p_e\approx 2 s_b/\xi^2$, for $p_e>s_m$),
so that the overall time to adaptation is robustly predicted to increase with offspring variance ($\xi^2$)
and decrease with population density and beneficial selection coefficient.
Furthermore, the probability that the second adaptation is a new mutation,
i.e.\ the probability of parallel adaptation, 
with $\gamma = \min(1,s_m/p_e)$,
is 
\begin{equation} \label{eqn:parallel_prob}
    \frac{\mutrate}{\mutrate+\migrate(R)} = \frac{ A \mu / (2s_m) }{ A \mu / (2s_m) + \gamma \left(R \sqrt{2 s_m} /\sigma \right)^{-(d-1)/2}\exp\left(- \sqrt{2 s_m} R / \sigma \right) },  
\end{equation}
so that the probability of parallel mutation should increase
approximately logistically with the distance between the patches, on the spatial scale $\sigma/\sqrt{s_m}$. \revpoint{2}{13}

\begin{figure}[ht!]
  \begin{center}
      % made in analyze-sims.R
      \includegraphics{prob-mutation-compared}
  \end{center}
  \caption{
      Probability that two patches separated by distance $R$
      evolve by independent mutations (i.e.\ in parallel),
      as predicted by theory (colors, contours)
      and in simulations (points).
      \textbf{With $C=5$.}
      XXX: more info
  }   \label{fig:sim_probs}
\end{figure}

>>>>

Taking these approximations at face value, 
and assuming both rates are small, 
the time until the first of the two patches is colonized by the $B$
allele will be approximately exponentially distributed with mean $1/(2
\mutrate)$.
Following this, the time until the second patch is subsequently colonized 
(via either migration or new mutation) 
will be exponentially distributed with mean $1/(\mutrate+\migrate)$.
Both these rates scale linearly with population density ($\rho$) 
and the probability of establishment of a rare allele ($p_e\approx 2 s_b/\xi^2$, for $p_e>s_m$),
so that the overall time to adaptation is robustly predicted to increase with offspring variance ($\xi^2$)
and decrease with population density and beneficial selection coefficient.
Furthermore, the probability that the second adaptation is a new mutation,
i.e.\ the probability of parallel adaptation, 
with $\gamma = \min(1,s_m/p_e)$,
is 
\begin{equation} \label{eqn:parallel_prob}
    \frac{\mutrate}{\mutrate+\migrate(R)} = \frac{ A \mu / (2s_m) }{ A \mu / (2s_m) + C \gamma \left(R \sqrt{2 s_m} /\sigma \right)^{-(d-1)/2}\exp\left(- \sqrt{2 s_m} R / \sigma \right) },  
\end{equation}
so that the probability of parallel mutation should increase
approximately logistically with the distance between the patches, on the spatial scale $\sigma/\sqrt{s_m}$. \revpoint{2}{13}

We tested this using the same simulations as figure \ref{fig:sim_times},
by using the empirical distributions of the respective times to adaptation
to estimate the probability,
for each parameter combination,
that a new, successful mutation appears 
before a successful migrant arrives from another, already adapted patch.
The results are compared to \eqref{eqn:parallel_prob}
in figure~\ref{fig:sim_probs}.

\begin{figure}[ht!]
  \begin{center}
      % made in analyze-sims.R
      \includegraphics{prob-mutation-compared}
  \end{center}
  \caption{
      Probability that two patches 
      evolve by independent mutations (i.e.\ in parallel),
      as a function of their geographic separation ($R$)
      and the strength of selection between patches ($s_m$).
      The background shades of grey and contours
      give the probability predicted by theory (equation~\eqref{eqn:parallel_prob}, with $C=5$).
      Points correspond to parameter combinations of the simulations in figure \ref{fig:sim_times},
      with the shade of grey and the red numeric label indicating the same probability
      estimated from these simulations (see text for details).
      The two densities, $\rho=100$ and $\rho=1000$ would lie directly on top of each other, so the $\rho=1000$ points have been shifted vertically.
  }   \label{fig:sim_probs}
\end{figure}
